---
title: "qPCR Experiment"
output: rmarkdown::html_vignette
package: qPCRExperiment
vignette: >
  %\VignetteIndexEntry{qPCR Experiment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Installation
```{r, echo=TRUE, eval=FALSE}
if(!require(qPCRExperiment)){
    BiocManager::install('NormqPCR')
    devtools::install_github('tjgorrie/miRoar')
}
library(qPCRExperiment)
```
```{r, eval=TRUE, echo=FALSE}
library(qPCRExperiment)
```

# Data Import
Importing data into R using the `qPCRExperiment` is very easy.
```{r, echo=TRUE, eval=TRUE}
input_directory <- system.file('extdata', 'edsfiles', package='qPCRExperiment')
example_experiment <- readEDSExperiment(input_directory, directory=TRUE, verbose=FALSE)
# As this example data contains two experiments with differing number of detectors we need to run
# the following steps, it is unlikely you would need to do this if you aren't mixing experiments as we are.
# Filter data to detectors present in all samples (in this case 6)
shared_detectors <- names(which(table(unlist(sapply(example_experiment[['plates']], detector_names))) == 6))
example_experiment <- example_experiment[shared_detectors,,]
print(example_experiment)
```

# The `Experiment` Object
The `Experiment` object attempts to collect together all the infomation from multiple EDS files. 
From the top-level down, the experiment contains a list of `Plate`s which are each individual EDS file that goes into
making the `Experiment` object. Within each `Plate` there are a list of `Well`s that contains the actual information from
the experiment. The data is organised in such a way that you should only ever need to interact with the `Experiment` and can 
manipulate the data as needed without ever changing the values inside the `Well`s

# Basic Processing

Experiment objects can be filtered using `[` methods to manipulate the number of samples, detectors or plates that 
we wish to analyse.
```{r, echo=TRUE, eval=FALSE}
example_experiment[detector_names,sample_names,plate_names]
```

```{r, echo=TRUE, eval=TRUE}
head(detector_names(example_experiment))
example_experiment[c("hsa-miR-937-3p_479212_mir", "hsa-miR-938_479214_mir", "hsa-miR-939-5p_478245_mir"),,]

sample_names(example_experiment)
example_experiment[,c("SALIVA-A", "SALIVA-B"),]

plate_names(example_experiment)
example_experiment[,,c("BXF34")]
```

# Examplar Quality Control
The `Experiment` object contains a couple of varieties cycle threshold values, Namely `Ct` and `Crt Raw`. These can be accessed via the `get` method:
```{r}
# The `deep=TRUE` argument lets the experiment object see data into the wells.
ct_vals <- example_experiment$get('Avg Ct', deep=TRUE)
plot(density(ct_vals))
crt_vals <- example_experiment$get('Crt Raw', deep=TRUE)
plot(density(crt_vals))
crt_amp <- example_experiment$get('Crt Amp', deep=TRUE)
plot(density(crt_amp))
raw_amp <- example_experiment$get('Raw Amp Score', deep=TRUE)
plot(density(raw_amp))
cq_conf <- example_experiment$get('Cq Conf', deep=TRUE)
plot(density(cq_conf))
```

For our analysis in xyz paper, we used a min `Cq Conf` of 0.8, A min `Crt Amp` of 0 and also limited our
analyses to 0 and 40.
Seeing as we extracted the raw data we can simply apply the filters to the objects in memory.
```{r}
crt_vals_filtered <- crt_vals
is.na(crt_vals_filtered) <- cq_conf < 0.8 | crt_amp < 0 | crt_vals <= 0 | crt_vals >= 40
dim(na.omit(crt_vals))
dim(na.omit(crt_vals_filtered))
```

# Calculating dCT Values
We provide a fancy method
```{r}
# Use global median to calculate dCt Values
global_dct <- deltaCt(crt_vals_filtered, method="global")

# Pick 1 or more HKs to use as factors (uses geometric mean of selected HKs)
endogenous_dct <- deltaCt(crt_vals_filtered, method="endogenous", HKs=c("hsa-miR-598-3p_478172_mir"))

# Perform geNorm method and use 10 HKs
genorm_dct <- deltaCt(crt_vals_filtered, method="geNorm", HKs=10)
# Perform NormFinder method and use 10 HKs, finding stable HKs between groups!
normfinder_dct <- deltaCt(crt_vals_filtered, method="NormFinder", HKs=10, group=c(1,2,2,1,1,2))

plot(density(na.omit(global_dct)))
lines(density(na.omit(endogenous_dct)), col=2)
lines(density(na.omit(genorm_dct)), col=3)
lines(density(na.omit(normfinder_dct)), col=4)
```

## FAQS

### Some of my plates have differing numbers of detectors how should I proceed
You can simply keep the plates with the incorrect number of detectors
```{r}
input_directory <- system.file('extdata', 'edsfiles', package='qPCRExperiment')
example_experiment <- readEDSExperiment(input_directory, directory=TRUE, verbose=FALSE)
print(sapply(example_experiment[['plates']], n_detectors))
plates_to_keep <- plate_names(example_experiment)[sapply(example_experiment[['plates']], n_detectors) == 754]
example_experiment[,,plates_to_keep]
```
or keep an intersection of the detectors present in all 
```{r}
print(head(table(unlist(sapply(example_experiment[['plates']], detector_names)))))
shared_detectors <- names(which(table(unlist(sapply(example_experiment[['plates']], detector_names))) == 6))
example_experiment[shared_detectors,,]
```